

# Peer Assessment 1 for Reproducible Research

https://github.com/snowdj/RepData_PeerAssessment1

## Loading and preprocessing the data

1.Load data comma-separated-value (CSV) file

```{r, echo = TRUE}
myData<- read.csv(unz("activity.zip","activity.csv"),header=T, sep=",")
# or myData <- read.csv("activity.csv",header=T)
```


Missing values are coded as NA

```{r, echo = TRUE}
myData<-na.omit(myData)
```

2.Process/transform the data (if necessary) into a format suitable for  analysis

```{r, echo = TRUE}
myData$steps <- as.numeric(myData$steps)
myData$date <- as.Date(myData$date, "%Y-%m-%d")
myData$interval <- as.numeric(myData$interval)
head(myData)
```

## What is mean total number of steps taken per day?

1.Make a histogram of the total number of steps taken each day

```{r, echo = TRUE}
steps_per_day <- aggregate(myData$steps, by=list(myData$date),sum)


str(steps_per_day)
list(myData$date)
hist(steps_per_day$x, , xlab="Steps_per_day")

tdf <- ddply(myData, ~ date, summarise, totalsteps=sum(steps, na.rm=TRUE))
hist(tdf$totalsteps, breaks=10, main="Histogram of Total Steps", xlab="Total Steps", ylab="Frequency")


```
2.Calculate and report the mean and median total number of steps taken per day

```{r, echo = TRUE}
mean(steps_per_day$x)

median(steps_per_day$x)

median(tdf$totalsteps)
mean(tdf$totalsteps)


```
## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo = TRUE}
steps_per_interval <- aggregate(myData$steps, by=list(myData$interval),mean)

plot(steps_per_interval$Group.1,steps_per_interval$x ,type = "l", xlab="Interval",ylab="Average number of steps take")


  idf <- ddply(df, ~ interval, summarise, meansteps=mean(steps, na.rm=TRUE))
  summary(idf)

  plot(idf$interval, idf$meansteps, type="l", main="", xlab="Interval", ylab="Average Steps")


```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo = TRUE}
steps_per_interval$Group.1[steps_per_interval$x==max(steps_per_interval$x)]
```
## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, echo = TRUE}
myData1<- read.csv(unz("activity.zip","activity.csv"),header=T, sep=",")
myData1$steps <- as.numeric(myData1$steps)
myData1$date <- as.Date(myData1$date, "%Y-%m-%d") # setup the format for the date
myData1$interval <- as.numeric(myData1$interval)
```

The total number of rows with NAs
```{r, echo = TRUE}
sum(is.na(myData1$steps))
```

2. Devise a strategy for filling in all of the missing values in the dataset with the mean for that 5-minute interval.

```{r, echo = TRUE}

# change the column names for steps_per_day and steps_per_interval


colnames(steps_per_interval)[1] <- "interval"

colnames(steps_per_interval)[2] <- "steps_per_interval"

```


3.Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo = TRUE }
# merger two 

myData1 <- merge(myData1,steps_per_interval,by="interval",all=TRUE)

myData1$steps[is.na(myData1$steps)] <- myData1$steps_per_interval[is.na(myData1$steps)]

  impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm=TRUE))
  dfimputed <- ddply(myData1, ~ interval, transform, steps = impute.mean(steps))
  
  #Sort the imputed data
  myData1imputed <- myData1[with(dfimputed, order(date, interval)), ]


```

4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r, echo = TRUE }
steps_per_day1 <- aggregate(myData1$steps, by=list(myData1$date),sum)

hist(steps_per_day1$x,breaks=10,xlab="Number of steps taken each day",main="")

  tdfimputed <- ddply(myData1imputed, ~ date, summarise, totalsteps=sum(steps, na.rm=TRUE))
  summary(tdf)
  hist(tdfimputed$totalsteps, breaks=10, main="Total Steps of Imputed Data", xlab="Total Steps", ylab="Frequency")

# the mean and median total number of steps taken per day.
mean(steps_per_day1$x)

median(steps_per_day1$x)

```

It has no impact.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo=TRUE}
myData1$weekday <- weekdays(myData1$date)
myData1$weekend<-myData1$weekday ==  "Sunday" | myData1$weekday ==  "Saturday" 
myData1$weekend <- factor(myData1$weekend, labels=c("weekday","weekend"))
```
2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

Using the xyplot to divide data to two parts, and plot the graphs. It shows therer are different patterns between weekday and wenkend. In weekday morning, people have more walk. In weekend during the day, people have more walk in general.  
```{r, echo=TRUE}
library(lattice)
xyplot(aggregate(myData1$steps, by=list(myData1$interval),mean)$x ~ aggregate(myData1$steps, by=list(myData1$interval),mean)$Group.1 | weekend, data = myData1, type = "l", layout = c(1, 2), xlab = "Interval",ylab = "Number of steps")
```